Machine learning high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2023-03-13 fame1_appendex.tex
fix fame1 appendix

@@

\section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
We can see the {\RAMBO} algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space. A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space. 
To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the {\RAMBO} mapping. As a result, each point on the resulting trajectory has equal probability density in phase-space. 
Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace {\RAMBO} in this procedure and could lead to trajectories with very different characteristics. One could for example imagine a 
sophisticated algorithm that only maps points close to the boundary of the hypercube to soft or collinear configurations. Using such an algorithm would have trajectories that avoid configurations with many particles more collinear or soft than 
the end-points of the trajectories. We find that with {\RAMBO} the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.

Figure~\ref{fig:detectorpath} shows the trajectory we chose in Section~\ref{sec:trajectories}.

Here we present another random phase-space trajectory similar to the one
in Figure~\ref{fig:random_trajectory} to demonstrate that features of that
figure are not unique.
\begin{figure}
    \includegraphics[width=\linewidth]{fame1/trajectory_1255.pdf}
    \caption{Another random phase-space trajectory where the `double+single'
    regions of phase-space are explicitly shown in blue.}
    \label{fig:trajectory_1255}
\end{figure}
\end{document}

>>>>

\section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
We can see the {\RAMBO} algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space. A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space. 
To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the {\RAMBO} mapping. As a result, each point on the resulting trajectory has equal probability density in phase-space. 
Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace {\RAMBO} in this procedure and could lead to trajectories with very different characteristics. One could for example imagine a 
sophisticated algorithm that only maps points close to the boundary of the hypercube to soft or collinear configurations. Using such an algorithm would have trajectories that avoid configurations with many particles more collinear or soft than 
the end-points of the trajectories. We find that with {\RAMBO} the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.
Figure~\ref{fig:detectorpath} shows the trajectory we chose in Section~\ref{sec:trajectories}.

@@

\section{Additional random phase-space trajectory}\label{appendix:fame1_trajectory2}
Here we present another random phase-space trajectory similar to the one
in Figure~\ref{fig:random_trajectory} to demonstrate that features of that
figure are not unique.
\begin{figure}[ht]
    \includegraphics[width=\linewidth]{fame1/trajectory_1255.pdf}
    \caption{Another random phase-space trajectory where the `double+single'
    regions of phase-space are explicitly shown in blue.}
    \label{fig:trajectory_1255}
\end{figure}

\end{document}

### 2023-03-13 paper1.tex
column wording
fix fame1 appendix

\subsection{Random trajectories}\label{sec:trajectories}
As another test of our emulator we assess the accuracy of our predictions on random phase-space trajectories.
These random phase-space trajectories are generated by connecting two random points in phase-space continuously without excluding any region of phase-space.
This presents an interesting and challenging test of the interpolation and extrapolation abilities of the NNs as some parts of the trajectories may lie outside of the phase-space region of the training datasets.
We show the results for 5-jet trajectories as this is the highest multiplicity we considered, predictions on lower multiplicities are better-behaved.
We investigated 50 different random trajectories. For the discussion in this section
we chose one that contains many interesting features,
namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
In Figure~\ref{fig:random_trajectory} we show the predictions by the three 5-jet models trained on data with different global phase-space cuts for this trajectory.
Left column shows the actual matrix element prediction and right column shows the ratios of the prediction to {\NJet}.
We analysed the predictions for the 50 random trajectories and measured the fraction of their length where the accuracy falls within given intervals.
Table~\ref{table:trajectory_predictions} shows the result for the regions of phase-space where training data was available, and those falling beyond the data available to the model.

@@

NN predictions are depicted as coloured scatter plots where the colour indicates the value of the minimum $s_{ij}$ between any pair of final-state particles at that phase-space point.
To more easily visualise the extrapolation performance of the NNs we highlight the regions where the minimum $s_{ij}$ goes below the global phase-space cut applied to the training set the models were trained on, 
for each cut made. The regions of the plots where this occurs have been coloured in red.  
With these trajectories being completely randomly selected in phase-space there is a possibility for there to be doubly singular points or worse.
To check for this we used {\FastJet} to cluster the phase-space points in the same way we did for data generation, see Section~\ref{sec:data}.
The pink regions indicate points which have two separate single unresolved limits, we label this configuration `Single+single'.
The purple regions indicate points which have a double unresolved limit (i.e. three particles in one jet), we label this configuration `Double'.
Although not seen in Figure~\ref{fig:random_trajectory}, there are points which have both a double unresolved limit and a separate single unresolved limit,
we label this configuration `Double+single' and indicate is as a blue region\footnote{Another random phase-space trajectory is presented in Appendix~\ref{appendix:fame1_trajectory_generation} which contains these `double+single' regions.}.
We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.
The pink, purple, and blue bands indicates regions of points which would have been discarded for our training and testing datasets.

>>>>

\subsection{Random trajectories}\label{sec:trajectories}
As another test of our emulator we assess the accuracy of our predictions on random phase-space trajectories.
These random phase-space trajectories are generated by connecting two random points in phase-space continuously without excluding any region of phase-space.
This presents an interesting and challenging test of the interpolation and extrapolation abilities of the NNs as some parts of the trajectories may lie outside of the phase-space region of the training datasets.
We show the results for 5-jet trajectories as this is the highest multiplicity we considered, predictions on lower multiplicities are better-behaved.
We investigated 50 different random trajectories. For the discussion in this section
we chose one that contains many interesting features,
namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
In Figure~\ref{fig:random_trajectory} we show the predictions by the three 5-jet models trained on data with different global phase-space cuts for this trajectory.
The left column shows the actual matrix element prediction and the right column shows the ratios of the prediction to {\NJet}.
We analysed the predictions for the 50 random trajectories and measured the fraction of their length where the accuracy falls within given intervals.
Table~\ref{table:trajectory_predictions} shows the result for the regions of phase-space where training data was available, and those falling beyond the data available to the model.

@@

NN predictions are depicted as coloured scatter plots where the colour indicates the value of the minimum $s_{ij}$ between any pair of final-state particles at that phase-space point.
To more easily visualise the extrapolation performance of the NNs we highlight the regions where the minimum $s_{ij}$ goes below the global phase-space cut applied to the training set the models were trained on, 
for each cut made. The regions of the plots where this occurs have been coloured in red.  
With these trajectories being completely randomly selected in phase-space there is a possibility for there to be doubly singular points or worse.
To check for this we used {\FastJet} to cluster the phase-space points in the same way we did for data generation, see Section~\ref{sec:data}.
The pink regions indicate points which have two separate single unresolved limits, we label this configuration `Single+single'.
The purple regions indicate points which have a double unresolved limit (i.e. three particles in one jet), we label this configuration `Double'.
Although not seen in Figure~\ref{fig:random_trajectory}, there are points which have both a double unresolved limit and a separate single unresolved limit,
we label this configuration `Double+single' and indicate is as a blue region\footnote{Another random phase-space trajectory is presented in Appendix~\ref{appendix:fame1_trajectory2} which contains these `double+single' regions.}.
We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.
The pink, purple, and blue bands indicates regions of points which would have been discarded for our training and testing datasets.
